---
title: "temoral_charge_points"
output:
  html_document:
    theme: yeti
    smart: true
    highlight: textmate
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---


## Load packages


### Load basic package

```{r}
library(spatstat)
library(here)
library(sp)
library(rgeos)
library(maptools)
library(GISTools)
library(tmap)
library(sf)
library(geojson)
library(geojsonio)
library(tmaptools)
library(tidyverse)
library(spdep)

```

### load plot library

```{r}
library(ggplot2)
library(plotly)
library(RColorBrewer)
```




## load data


### load .csv file


```{r}
charge <- read_csv("national-charge-point-registry.csv")
summary(charge$chargeDeviceStatus)
```


### load shapefile

```{r}
londonwards <- st_read("ESRI/London_Ward_CityMerged.shp") %>% 
  st_transform(., 27700)
```
```{r}
london_outline <- londonwards %>%
  summarise(area = sum(HECTARES))
```




## filter data



```{r}
chargepoints <- charge [charge$chargeDeviceStatus=="In service" & !is.na(charge$latitude) & !is.na(charge$longitude),] 
```


### function to filter year (2014, 2017, 2020 register)

```{r}
filterfun <- function(end){
  enddate <- end
  
  filterpoints <- chargepoints %>% 
    filter(chargepoints$dateCreated<=enddate)
  return(filterpoints)
}

```


### function to transform projection & crop to London

```{r}
londonPfun <- function(filteredpoints){
  transpoints <- filteredpoints %>% 
    select(., c(3,4,5)) %>% 
    st_as_sf(., coords = c("longitude", "latitude"),crs = 4326) %>% 
    st_transform(., 27700) %>%
    distinct()
  
  charge_london <- st_intersects(london_outline, transpoints)
  londonpoints <- transpoints[unlist(charge_london),]
  return(londonpoints)
}
```


### instance object

```{r}
filterpoints14 <- filterfun("2014-12-31")
londonpoints2014 <- londonPfun(filterpoints14)

filterpoints17 <- filterfun("2017-12-31")
londonpoints2017 <- londonPfun(filterpoints17)

filterpoints20 <- filterfun("2020-12-31")
londonpoints2020 <- londonPfun(filterpoints20)
```


### plot data

```{r}
tmap_mode("plot")

tm1 <- tm_shape(london_outline) +
  tm_fill() +
  tm_shape(londonpoints2014) +
  tm_dots(col = "red", size = .01)+
  tm_layout(frame=FALSE)+
  tm_credits("(a)", position=c(0,0.85), size=1)

tm2 <- tm_shape(london_outline) +
  tm_fill() +
  tm_shape(londonpoints2017) +
  tm_dots(col = "red", size = .01)+
  tm_layout(frame=FALSE)+
  tm_credits("(b)", position=c(0,0.85), size=1)

tm3 <- tm_shape(london_outline) +
  tm_fill() +
  tm_shape(londonpoints2020) +
  tm_dots(col = "red", size = .01)+
  tm_layout(frame=FALSE)+
  tm_credits("(c)", position=c(0,0.85), size=1)


t <- tmap_arrange(tm1, tm2, tm3, ncol=3)
t

```
```{r}
tmap_save(t, 'london charge points.png')
```




## plot density hex


### get coordinates from sf object

```{r}
# get coordinates from sf object -> ggplot2 format
coordsfun <- function(londonpoints, year){
  charge_points_coords <- matrix(unlist(londonpoints$geometry),ncol=2,byrow=T) %>%
    as.data.frame() %>% 
    na.omit() %>% 
    mutate(year=year)
  return(charge_points_coords)
}

# inspect
charge_points_coords14 <- coordsfun(londonpoints2014, "2014")
head(charge_points_coords14)
```


### see trend in massive datasets

```{r}
charge_points_coords14 <- coordsfun(londonpoints2014, "2014")
charge_points_coords17 <- coordsfun(londonpoints2017, "2017")
charge_points_coords20 <- coordsfun(londonpoints2020, "2020")

charge_points_ggplot <- rbind(charge_points_coords14, 
                              charge_points_coords17, 
                              charge_points_coords20)

```

```{r}
class(london_outline)
# methods(class = "sf")
df <- as.data.frame(st_coordinates(london_outline))
class(df)
```
```{r}
ggplot()+
  geom_polygon(data=df, aes(X, Y, group=L1))+
  coord_quickmap()
```


```{r}

ggplot(charge_points_ggplot, aes(x=V1, y=V2, group=year)) +
  geom_hex(bins=50, na.rm=TRUE) +
  geom_polygon(data=df, fill=NA, 
               color="black", size=.2,
               aes(X, Y, group=L1)) +
  labs(fill = "charge point count \n", 
       title = "Charge Points in London \n") +
  scale_fill_continuous(high=brewer.pal(7, "Reds")[7], 
                        low=brewer.pal(7, "Reds")[1]) +
  xlab(NULL)+
  ylab(NULL)+
  scale_x_discrete()+
  scale_y_discrete()+
  scale_colour_gradient()+
  facet_wrap(~ year, ncol = 2)+
  theme_minimal()+
  theme(strip.text = element_text(size=11), 
        legend.key.size = unit(11, "pt"),
        legend.text = element_text(size=8), 
        legend.title = element_text(size=11))+
  coord_quickmap()

```
```{r}
ggsave("charge points facet wrap.jpg")
```




## Ripley's K



```{r}
# transform sf to sp
london_sp <- as(londonwards,'Spatial')
# get the window of observation using maptools package
window <- as.owin.SpatialPolygons(london_sp)
# inspect
window

```

```{r}
# create ppp object
charge_ppp <- ppp(x=charge_points_coords20[,1],
                  y=charge_points_coords20[,2],
                  window=window,check=T)

# inspect
plot(charge_ppp)
```
```{r}
# check for duplicates
any(duplicated(charge_ppp))

# count the number of duplicated points
sum(multiplicity(charge_ppp) > 1)

```
```{r}
# get the average distance to the first, second, ..., the hundredth, nearest neighbour
charge_ann <- apply(nndist(charge_ppp, k=1:100),2,FUN=mean)
# plot the results
plot(charge_ann ~ seq(1:100))

```
```{r}
# calculate Ripley's K for charge locations, maximum radius of 4 kilometers
plot(Kest(charge_ppp,correction='border',rmax=2000))
```



## density analysis



```{r}
# create and plot a completely spatially random point pattern of the same size as charge points data
plot(rpoispp(3817))

```
```{r}
# chi-square between observed pattern and Poisson sampled points
quadrat.test(charge_ppp,nx=15,ny=15)
```


### kernel density analysis

```{r}
# kernel density estimation with a 1000 metre bandwidth
plot(density.ppp(charge_ppp,sigma=1000))
```

### dbscan clustering analysis

```{r}
library(fpc)
```

```{r}
# Run dbscan's DB-Scan function with eps and minPts set as follows
chargepoints_dbscan <- fpc::dbscan(charge_points_coords20[1:2], eps = 3000, MinPts = 200)

# Plot our clusters and points - note the plotting code for the dbscan plot is different to that of fpc
plot(chargepoints_dbscan, 
     londonpoints2020, 
     main = "DBSCAN for charge points")
plot(london_outline$geometry, add=T)
```

### k nearest neighbor distance

```{r}
library(dbscan)
charge_points_coords20[1:2]%>%
  dbscan::kNNdistplot(.,k=4)
```


### function to map charge point density by wards


```{r}
pointsjoinedfun <- function(londonpointsyear){
  londonpointsyear <- londonpointsyear
  charge_points_ward <- londonpointsyear[londonwards,]
  
  points_wards <- londonwards%>%
    st_join(charge_points_ward)%>%
    add_count(NAME)%>%
    janitor::clean_names()%>%
    #then density of the points per ward
    mutate(density=(n/hectares)*1000) %>% 
    mutate(id=1) %>% 
    mutate(sumpoints=aggregate(n~id,.,FUN=sum)) %>% 
    mutate(percentage=((n/sumpoints[1, 2]))*100) %>% 
    #select density and some other variables 
    dplyr::select(density, percentage, name, gss_code, n) %>% 
    distinct(., gss_code, .keep_all = TRUE)
  
  # same quick choropleth map
  points_wards <- points_wards %>%                    
    group_by(gss_code) %>%         
    summarise(density=first(density),
              percentage=first(percentage),
              wardname=first(name),
              chargepointcount= first(n))
  return(points_wards)
}
```
```{r}
# test
percent_ward14 <- pointsjoinedfun(londonpoints2014)
head(percent_ward14)
```


```{r}
densityfun <- function(londonpointsyear, number){
  points_wards <- pointsjoinedfun(londonpointsyear)

  breaks = c(0, 20, 30, 40, 50, 60, 200, 400, +Inf) 
  tm <- tm_shape(points_wards) +
    tm_polygons("density",
                breaks=breaks,
                palette=RColorBrewer::brewer.pal(8, "YlOrRd"),
                midpoint=NA) +
    tm_legend(show=FALSE)+
    tm_layout(frame=FALSE)+
    tm_credits(number, position=c(0,0.85), size=1)
  
  return(tm)
}

```
```{r}
legendfun <- function(londonpointsyear){
  points_wards <- pointsjoinedfun(londonpointsyear)
  
  breaks = c(0, 20, 30, 40, 50, 60, 200, 400, +Inf) 
  legend <- tm_shape(points_wards) +
    tm_polygons("density",
                breaks=breaks,
                palette=RColorBrewer::brewer.pal(8, "YlOrRd"), 
                title="Density of Charge Points in London \n(per thousand hectare)") +
    tm_scale_bar(position=c(0.2,0.04), text.size=0.6)+
    tm_compass(north=0, position=c(0.65,0.2))+
   
    tm_layout(title = "Charge Points Density Trend", 
              legend.title.size=1,
              legend.text.size = 0.6,
              legend.only = TRUE, 
              legend.position=c(0.2,0.25),asp=0.1)
  
  return(legend)
}
```


### tmap - charge point density by ward

```{r}

tm1 <- densityfun(londonpoints2014, "a)")
tm2 <- densityfun(londonpoints2017, "b)")
tm3 <- densityfun(londonpoints2020, "c)")

legend <- legendfun(londonpoints2020)


t_density <- tmap_arrange(tm1, tm2, tm3, legend, ncol=2)
t_density
```

```{r}
tmap_save(t_density, 'london charge points density.png')
```


### function to map charge point percentage by wards

```{r}
percentfun <- function(londonpointsyear, number){
  points_wards <- pointsjoinedfun(londonpointsyear)

  breaks = c(0, 0.05, 0.1, 0.15, 0.2, 0.3, 0.5, 0.7, +Inf) 
  tm <- tm_shape(points_wards) +
    tm_polygons("percentage",
                breaks=breaks,
                palette=RColorBrewer::brewer.pal(8, "YlOrRd"),
                midpoint=NA) +
    tm_legend(show=FALSE)+
    tm_layout(frame=FALSE)+
    tm_credits(number, position=c(0,0.85), size=1)
  
  return(tm)
}

```
```{r}
legendpfun <- function(londonpointsyear){
  points_wards <- pointsjoinedfun(londonpointsyear)
  
  breaks = c(0, 0.05, 0.1, 0.15, 0.2, 0.3, 0.5, 0.7, +Inf)  
  legend <- tm_shape(points_wards) +
    tm_polygons("percentage",
                breaks=breaks,
                palette=RColorBrewer::brewer.pal(8, "YlOrRd"), 
                title="Percentage of Charge Points in London \n(by wards)") +
    tm_scale_bar(position=c(0.2,0.04), text.size=0.6)+
    tm_compass(north=0, position=c(0.65,0.2))+
   
    tm_layout(title = "Charge Points Distribution Trend", 
              legend.title.size=1,
              legend.text.size = 0.6,
              legend.only = TRUE, 
              legend.position=c(0.2,0.25),asp=0.1)
  
  return(legend)
}
```


### tmap - charge point density by ward

```{r}

p1 <- percentfun(londonpoints2014, "a)")
p2 <- percentfun(londonpoints2017, "b)")
p3 <- percentfun(londonpoints2020, "c)")

legendp <- legendpfun(londonpoints2020)


percentage <- tmap_arrange(p1, p2, p3, legendp, ncol=2)
percentage
```
```{r}
tmap_save(percentage, 'Charge Points Distribution Trend.png')
```




## spatial autocorrelation


### function to calculate the centroids

```{r}
# library(spdep)
# First calculate the centroids of all Wards in London

coordsWfun <- function(londonpointsyear){
  points_wards <- pointsjoinedfun(londonpointsyear)
  coordsW <- points_wards %>% 
    st_centroid() %>%
    st_geometry()
}

```
```{r}
# test
coordsW <- coordsWfun(londonpoints2020)
plot(coordsW,axes=TRUE)
```


### function to create a neighbours list

```{r}
#create a neighbours list
ward_nbfun <- function(londonpointsyear){
  points_wards <- pointsjoinedfun(londonpointsyear)
  ward_nb <- points_wards %>% 
    poly2nb(., queen=T)
  return(ward_nb)
}

```
```{r}
# test
ward_nb20 <- ward_nbfun(londonpoints2020)
plot(ward_nb20, st_geometry(coordsW), col="red")
# add a map underneath
points_wards <- pointsjoinedfun(londonpoints2020) 
plot(points_wards$geometry, add=T)
```

### function to create spatial weights

```{r}
#create a spatial weights object from these weights
ward_lwfun <- function(londonpointsyear){
  ward_lw <- ward_nbfun(londonpointsyear) %>% 
    nb2listw(., style="C")
  return(ward_lw)
}
```
```{r}
# test
ward_lw20 <- ward_lwfun(londonpoints2020)
head(ward_lw20$neighbours)
```


### function to run Global Moran's I

```{r}
I_ward_globalfun <- function(londonpointsyear){
  ward_lw <- ward_lwfun(londonpointsyear)
  
  I_ward_global <- londonpointsyear %>% 
    pointsjoinedfun(.) %>% 
    pull(density) %>%
    as.vector()%>% # Converts a distributed matrix into a non-distributed vector
    moran.test(., ward_lw) # density <-> spatial (similar or not)
  return(I_ward_global)
}

```


### Global Moran's I for 2014, 2017, 2020

```{r}
I_ward_global14 <- I_ward_globalfun(londonpoints2014)
I_ward_global17 <- I_ward_globalfun(londonpoints2017)
I_ward_global20 <- I_ward_globalfun(londonpoints2020)
I_ward_global14
I_ward_global17
I_ward_global20
```


### function to run Local Moran's I

```{r}
ward_lw20 <- ward_lwfun(londonpoints2020)
  
  I_ward_local_count <- londonpoints2020 %>% 
    pointsjoinedfun(.) %>% 
    pull(chargepointcount) %>%
    as.vector()%>% 
    localmoran(., ward_lw20) %>% 
    as_tibble()
  I_ward_local_density <- londonpoints2020 %>% 
    pointsjoinedfun(.) %>% 
    pull(density) %>%
    as.vector()%>% 
    localmoran(., ward_lw20) %>% 
    as_tibble()

slice_head(I_ward_local_count, n=5)
slice_head(I_ward_local_density, n=5)
```


```{r}
# use the local moran function to generate I for each ward in the city

I_ward_localfun <- function(londonpointsyear){
  ward_lw <- ward_lwfun(londonpointsyear)
  
  I_ward_local_count <- londonpointsyear %>% 
    pointsjoinedfun(.) %>% 
    pull(chargepointcount) %>%
    as.vector()%>% 
    localmoran(., ward_lw) %>% 
    as_tibble()
  I_ward_local_density <- londonpointsyear %>% 
    pointsjoinedfun(.) %>% 
    pull(density) %>%
    as.vector()%>% 
    localmoran(., ward_lw) %>% 
    as_tibble()
  
  tm_local_moran <- londonpointsyear %>% 
    pointsjoinedfun(.) %>%
    mutate(charge_count_I = as.numeric(I_ward_local_count$Ii))%>%
    mutate(charge_count_Iz =as.numeric(I_ward_local_count$Z.Ii))%>%
    mutate(density_I =as.numeric(I_ward_local_density$Ii))%>%
    mutate(density_Iz =as.numeric(I_ward_local_density$Z.Ii))
  
  return(tm_local_moran)
}

```


### function to run tmap for Local Moran's I

```{r}
tmap_local_moranfun <- function(londonpointsyear, number){
  breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)
  MoranColours<- rev(brewer.pal(8, "RdBu"))
  tm_local_moran <- I_ward_localfun(londonpointsyear)
  
  tmap <- tm_shape(tm_local_moran) +
    tm_polygons("charge_count_Iz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA)+
    tm_legend(show=FALSE)+
    tm_layout(frame=FALSE)+
    tm_credits(number, position=c(0,0.85), size=1)
  return(tmap)
}
```
```{r}
tmap_moranlegendfun <- function(londonpointsyear){
  breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)
  MoranColours<- rev(brewer.pal(8, "RdBu"))
  tm_local_moran <- I_ward_localfun(londonpointsyear)
  legend <- tm_shape(tm_local_moran) +
    tm_polygons("charge_count_Iz",
                breaks=breaks1,
                palette=MoranColours,
                title="Local Moran's I, Charge Points in London") +
    tm_scale_bar(position=c(0.2,0.04), text.size=0.6)+
    tm_compass(north=0, position=c(0.65,0.2))+
   
    tm_layout(title = "Charge Points Local Moran's I", 
              legend.title.size=1,
              legend.text.size = 0.6,
              legend.only = TRUE, 
              legend.position=c(0.2,0.25),asp=0.1)
  
  return(legend)
}
```


```{r}
tm_local_moran14 <- tmap_local_moranfun(londonpoints2014, "a)")
tm_local_moran17 <- tmap_local_moranfun(londonpoints2017, "b)")
tm_local_moran20 <- tmap_local_moranfun(londonpoints2020, "c)")
```
```{r}
moran_legend <- tmap_moranlegendfun(londonpoints2020)

local_moran <- tmap_arrange(tm_local_moran14, 
                            tm_local_moran17, 
                            tm_local_moran20, 
                            moran_legend, ncol=2)
local_moran
```

```{r}
tmap_save(local_moran, "Local Moran's I, Charge Points in London.png")
```


### function to run Gi*

```{r}
Gi_ward_local_densityfun <- function(londonpointsyear){
  ward_lw <- ward_lwfun(londonpointsyear)
  
  Gi_ward_local_density <- londonpointsyear %>% 
    pointsjoinedfun(.) %>% 
    pull(density) %>%
    as.vector()%>%
    localG(., ward_lw)
  
  density_Gi <- londonpointsyear %>% 
    pointsjoinedfun(.) %>% 
    mutate(density_G = as.numeric(Gi_ward_local_density))
  
  return(density_Gi)
}
```


### function to run tmap for Gi*

```{r}
tmap_Gifun <- function(londonpointsyear, number){
  breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)
  GIColours<- rev(brewer.pal(8, "RdBu"))
  density_Gi <- Gi_ward_local_densityfun(londonpointsyear)
  
  tmap <- tm_shape(density_Gi) +
    tm_polygons("density_G",
                breaks=breaks1,
                palette=GIColours,
                title="Gi*, Charge Points in London") +
    tm_legend(show=FALSE)+
    tm_layout(frame=FALSE)+
    tm_credits(number, position=c(0,0.85), size=1)
  
  return(tmap)
}
```

```{r}
G1 <- tmap_Gifun(londonpoints2014, "a)")
G2 <- tmap_Gifun(londonpoints2017, "b)")
G3 <- tmap_Gifun(londonpoints2020, "c)")
```

```{r}
tmap_Gilegendfun <- function(londonpointsyear){
  breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)
  GIColours<- rev(brewer.pal(8, "RdBu"))
  density_Gi <- Gi_ward_local_densityfun(londonpointsyear)
  legend <- tm_shape(density_Gi) +
    tm_polygons("density_G",
                breaks=breaks1,
                palette=GIColours,
                title="Gi*, Charge Points in London") +
    tm_scale_bar(position=c(0.2,0.04), text.size=0.6)+
    tm_compass(north=0, position=c(0.65,0.2))+
   
    tm_layout(title = "Charge Points Gi* Density", 
              legend.title.size=1,
              legend.text.size = 0.6,
              legend.only = TRUE, 
              legend.position=c(0.2,0.25),asp=0.1)
  
  return(legend)
}
```
```{r}
Gi_legend <- tmap_Gilegendfun(londonpoints2020)

Gi <- tmap_arrange(G1, G2, G3, Gi_legend, ncol=2)
Gi
```

